---
title: "Project1 Part2 Applied Statistics"
author: "Abhina Premachandran"
date: "2023-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monte Carlo Simulations of DIS and HPQ stocks
 This project uses Monte Carlo Simulations to simulate daily returns of an investment portfolio. The simulated return mean is compared with actual return mean. The chosen stocks for this project are 'DIS' and 'HPQ' stocks. Further an optimal portfolio is suggested based on three scenarios on investing the amount on DIS and HPQ stocks.

```{r echo=FALSE}
suppressWarnings(suppressMessages(library(quantmod)))
```

Getting the stock data using `getSymbols`. 

```{r}
getSymbols(c('DIS','HPQ','USO','DVN','LMT'), from = Sys.Date() - 150, to = Sys.Date()) 
```
plotting to see the timeseries of DIS and HPQ stocks for the past 150 days
```{r fig.height=3.5, fig.fullwidth=TRUE}
par(mfrow = c(1,2))
plot(DIS$DIS.Close, main = 'DIS',type = 'l',col = 'blue', lwd = 2)
plot(HPQ$HPQ.Close, main = 'HPQ',type = 'l',col = 'violet', lwd = 2)
plot(USO$USO.Close, main = 'USO',type = 'l',col = 'green', lwd = 2)
plot(DVN$DVN.Close, main = 'DVN',type = 'l',col = 'red', lwd = 2)
plot(LMT$LMT.Close, main = 'DIS',type = 'l',col = 'cyan', lwd = 2)
```

Defining the stock price matrix for each stocks

```{r}
#DIS_Price <- as.matrix(as.numeric(DIS$DIS.Close)) 
#HPQ_Price <- as.matrix(as.numeric(HPQ$HPQ.Close)) 

#for( i in c(DIS,HPQ,USO,DVN,LMT)){
  #Price <- as.matrix(as.numeric(i$i.Close))
  #paste(d,Price)
#}
```


## computing the daily returns using diff()/lag() of Closing prices

```{r}
DIS$DIS.Returns <- diff(DIS$DIS.Close)/lag(DIS$DIS.Close)
Returns_DIS <- DIS$DIS.Returns
```
```{r}
HPQ$HPQ.Returns <- diff(HPQ$HPQ.Close)/lag(HPQ$HPQ.Close)
Returns_HPQ <- HPQ$HPQ.Returns
#Defining N_days as the length of the time series
N_days <- length(Returns_HPQ)
```
## Plotting the daily returns

```{r}
par(mfrow = c(1, 2))
hist(Returns_DIS,main = 'DIS', xlab = 'Daily Returns', col='blue',breaks = 20, freq=TRUE)
hist(Returns_HPQ,main = 'HPQ', xlab = 'Daily Returns', col='violet',breaks = 20, freq=TRUE)
```
From the histogram of each stock, it is clear that the daily returns of the stocks approximately follows a normal distribution. Therefore rnorm() can be used to simulate future returns.
## Computing the mean and standard deviation from the past stock returns inorder to simulate future returns 
I am choosing the returns upto 30 days before as the past data to simulate the rest of the 30 days(future stock) until today
```{r}
#To simulate for 30 days until today

DIS_daily_mean <- mean(Returns_DIS[(N_days-31):N_days])
DIS_daily_std <- sd(Returns_DIS[(N_days-31):N_days])

HPQ_daily_mean <- mean(Returns_HPQ[(N_days-31):N_days])
HPQ_daily_std <- sd(Returns_HPQ[(N_days-31):N_days])
```

## Monte Carlo Simulation for DIS
rnorm() is used to simulate returns and price for 30 days until today for DIS stock.
```{r}
set.seed(101)
DIS_starting_price <- DIS$DIS.Close[[(N_days-30)]]
returns <- 1+rnorm(30,mean=DIS_daily_mean,sd=DIS_daily_std)
no_sims <- 1001
return_list <- matrix(0, nrow=no_sims,ncol=30)
price_list <- matrix(0,nrow=no_sims,ncol=31)
for(i in 1:no_sims){
  return_list[i,]<-rnorm(30,mean=DIS_daily_mean,sd=DIS_daily_std)
  price_list[i,]<-cumprod(c(DIS_starting_price,1+return_list[i,]))
}
total_returns1 <- array(NA,dim=no_sims,dimnames=NULL)
for(i in 1:no_sims){
  total_returns1[i]<-(price_list[i,31]-price_list[i,1])/price_list[i,1]
}
```
## Monte Carlo Simulation for HPQ

```{r}
set.seed(101)
HPQ_starting_price <- HPQ$HPQ.Close[[(N_days-30)]]
returns <- 1+rnorm(30,mean=HPQ_daily_mean,sd=HPQ_daily_std)
no_sims <- 1001
return_list <- matrix(0, nrow=no_sims,ncol=30)
price_list <- matrix(0,nrow=no_sims,ncol=31)
for(i in 1:no_sims){
  return_list[i,]<-rnorm(30,mean=HPQ_daily_mean,sd=HPQ_daily_std)
  price_list[i,]<-cumprod(c(HPQ_starting_price,1+return_list[i,]))
}
total_returns2 <- array(NA,dim=no_sims,dimnames=NULL)
for(i in 1:no_sims){
  total_returns2[i]<-(price_list[i,31]-price_list[i,1])/price_list[i,1]
}

```


## Building a portfolio
We will assume an Initial Investment of $10,000 and allocate our funds between these two stocks based on 3 strategies:

- The entire investment to DIS
- The entire investment to HPQ
- Split the investment evenly between DIS and HPQ

## Calculating the mean and standard deviations of DIS and 
```{r}
mu1 <- mean(total_returns1)
sigma1 <- sd(total_returns1)
mu2 <- mean(total_returns2)
sigma2 <- sd(total_returns2)
ClosePrices <- merge(DIS$DIS.Close,HPQ$HPQ.Close)
rho_12 <- cor(ClosePrices, use = "pairwise.complete.obs")

total_return <- function(w){
  return(w*mu1 + (1-w)*mu2)
}
total_risk <- function(w){
  return(w^2*sigma1^2 + (1-w)^2*sigma2^2 + 2*w*(1-w)*rho_12*sigma1*sigma2)
}


```
From the previous computations, the cost of DIS is \$`r round(DIS_starting_price,2)`. Our analysis has shown that their expected return is \$`r round(DIS_mu, digits=4)` per share with a standard deviation of \$`r round(DIS_sigma, digits=4)`. Shares of HPQ cost \$`r round(HPQ_starting_price,2)` per share, with an expected return of \$`r round(HPQ_mu, digits=4)` and a standard deviation of \$`r round(HPQ_sigma, digits=4)`. We will assume that the returns from the two companies are independent. 

- 1) The entire investment to DIS
  For entire investment to DIS, w = 1. The total return is given by \$`total_return(1)` and the total risk is given by \$`total_risk(1)`

```{r}
# returns <- NULL
# risks <- NULL

# returns <- c(returns, total_return(w))
# risks <- c(risks, total_risk(w))
# 
# min_risk_w <- weights[which(risks == min(risks))]
# min_risk_return <- mean(cal_pf_return(min_risk_w))
# 
# efficient_idx <- which(returns >= min_risk_return)
# inefficient_idx <- which(returns < min_risk_return)
# 
# plot(risks[efficient_idx], returns[efficient_idx],
#      xlab = "Portfolio risk",
#      ylab = "Portfolio return",
#      main = "The Efficient Frontier",
#      ylim = c(0,0.2), xlim = c(0.002, 0.01))
# 
# lines(risks[inefficient_idx], returns[inefficient_idx], lty = "dotted")
# 
# abline(h = min_risk_return)
# text(x = 0.003, y = min_risk_return-0.01, "min variance portfolio")
# text(x = 0.006, y = 0.15, "efficient frontier")
```

```{r,echo = FALSE}
# lines for the next part of the Project 1

 plot(as.numeric(ClosePrices[,1]),as.numeric(ClosePrices[,2]),pch=20, xlab='DIS', ylab='HPQ')
```