---
title: "Project1 Part1 Applied Statistics"
author: "Abhina Premachandran"
date: "2023-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monte Carlo Simulations of DIS and HPQ stocks
 This project uses Monte Carlo Simulations to simulate daily returns of an investment portfolio. The simulated return mean is compared with actual return mean. The chosen stocks for this project are 'DIS' and 'HPQ' stocks. Further an optimal portfolio is suggested based on three scenarios on investing the amount on DIS and HPQ stocks.

```{r echo=FALSE}
suppressWarnings(suppressMessages(library(quantmod)))
```

Getting the stock data using `getSymbols`. 

```{r}
getSymbols(c('DIS','HPQ'), from = Sys.Date() - 150, to = Sys.Date()) 
```
plotting to see the timeseries of DIS and HPQ stocks for the past 150 days
```{r}
par(mfrow = c(1,2))
plot(DIS$DIS.Close, main = 'DIS',type = 'l',col = 'blue', lwd = 2)
plot(HPQ$HPQ.Close, main = 'HPQ',type = 'l',col = 'violet', lwd = 2)
```

Defining the stock price matrix for each stocks

```{r}
DIS_Price <- as.matrix(as.numeric(DIS$DIS.Close)) 
HPQ_Price <- as.matrix(as.numeric(HPQ$HPQ.Close)) 
```


## computing the daily returns using diff()/lag() of Closing prices

```{r}

DIS$DIS.Returns <- diff(DIS$DIS.Close)/lag(DIS$DIS.Close)
Returns_DIS <- DIS$DIS.Returns
```
```{r}
HPQ$HPQ.Returns <- diff(HPQ$HPQ.Close)/lag(HPQ$HPQ.Close)
Returns_HPQ <- HPQ$HPQ.Returns
#Defining N_days as the length of the time series
N_days <- length(Returns_HPQ)
```
## Plotting the daily returns

```{r}
par(mfrow = c(1, 2))
hist(Returns_DIS,main = 'DIS', xlab = 'Daily Returns', col='blue',breaks = 20, freq=TRUE)
hist(Returns_HPQ,main = 'HPQ', xlab = 'Daily Returns', col='violet',breaks = 20, freq=TRUE)
```
From the histogram of each stock, it is clear that the daily returns of the stocks approximately follows a normal distribution. Therefore rnorm() can be used to simulate future returns.
## Computing the mean and standard deviation from the past stock returns inorder to simulate future returns 
I am choosing the returns upto 30 days before as the past data to simulate the rest of the 30 days(future stock) until today
```{r}
#To simulate for 30 days until today

DIS_daily_mean <- mean(Returns_DIS[(N_days-31):N_days])
DIS_daily_std <- sd(Returns_DIS[(N_days-31):N_days])

HPQ_daily_mean <- mean(Returns_HPQ[(N_days-31):N_days])
HPQ_daily_std <- sd(Returns_HPQ[(N_days-31):N_days])
```

## Monte Carlo Simulation for DIS
rnorm() is used to simulate returns and price for 30 days until today for DIS stock.
```{r}
set.seed(101)
DIS_starting_price <- DIS$DIS.Close[[(N_days-30)]]
returns <- 1+rnorm(30,mean=DIS_daily_mean,sd=DIS_daily_std)
no_sims <- 1001
return_list <- matrix(0, nrow=no_sims,ncol=30)
price_list <- matrix(0,nrow=no_sims,ncol=31)
for(i in 1:no_sims){
  return_list[i,]<-rnorm(30,mean=DIS_daily_mean,sd=DIS_daily_std)
  price_list[i,]<-cumprod(c(DIS_starting_price,1+return_list[i,]))
}
total_returns1 <- array(NA,dim=no_sims,dimnames=NULL)
for(i in 1:no_sims){
  total_returns1[i]<-(price_list[i,31]-price_list[i,1])/price_list[i,1]
}

```
## Analysing and visualizing the simulations for DIS


```{r fig.height=3.5, fig.fullwidth=TRUE}

max1<-which.max(total_returns1)
min1<-which.min(total_returns1)
actual_prices_DIS <- as.matrix(as.numeric(DIS$DIS.Close[(N_days-30):N_days]))
med1<-match(median(total_returns1),total_returns1)

par(mfrow = c(1,2))
lb <- min(price_list[min1,])
ub <- max(price_list[max1,])

#Plotting the min, max, median of the simulated return along with the actual return
plot(0:30, price_list[min1,], type='l', main = 'Simulated price of DIS',ylab='Price',xlab='Days',ylim=c(lb,ub),col="red")
lines(0:30, price_list[med1,],type='l',col='black')
lines(0:30, price_list[max1,],type='l',col='green')
lines(0:30, actual_prices_DIS,type='l',col='blue')
grid(lty = 2, col = 'gray', lwd = 1)

hist(total_returns1, main = 'DIS', xlab = 'Simulated Total Returns', col='blue', freq=TRUE)
actual_return1=(DIS$DIS.Close[[N_days]]-DIS$DIS.Close[[(N_days-30)]])/DIS$DIS.Close[[(N_days-30)]]
abline(v=actual_return1, col = 'red', lwd=2)
abline(v=0, col = 'gray', lty=2, lwd=4)
```
In the plot, Simulations leading to the largest gain and largest loss in total 30-day return are shown in red and green respectively. The median from all performed simulations is shown in black and the actual returns are shown in blue. The histogram shows the distributions of the total returns as well as the actual 30-day return shown in red for comparison. 
## Comparing DIS stock simulations with the actual return
```{r}
DIS_expected_return <- mean(total_returns1)
DIS_return_std <- sd(total_returns1)

summary(total_returns1)
actual_return1

```

The summary statistics from all Monte Carlo trials had an average of `r mean(total_returns1)`. The actual return is `r actual_return1`. The actual and simulated returns for DIS stocks are not close.

## Estimating the probability of gaining or losing more than 10% return during the selected time period for DIS
```{r}

DIS_prob_G<-sum(total_returns1 > 0.1)/no_sims
DIS_prob_L<-sum(total_returns1 < (-0.1))/no_sims
```
As per our simulation, the estimated probability to gain more than 10% in 10 days for Disney is `r DIS_prob_G*100`% and the probability to lose more than 10% in 10 days is `r DIS_prob_L*100`.

## Monte Carlo Simulation for HPQ

```{r}
set.seed(101)
HPQ_starting_price <- HPQ$HPQ.Close[[(N_days-30)]]
returns <- 1+rnorm(30,mean=HPQ_daily_mean,sd=HPQ_daily_std)
no_sims <- 1001
return_list <- matrix(0, nrow=no_sims,ncol=30)
price_list <- matrix(0,nrow=no_sims,ncol=31)
for(i in 1:no_sims){
  return_list[i,]<-rnorm(30,mean=HPQ_daily_mean,sd=HPQ_daily_std)
  price_list[i,]<-cumprod(c(HPQ_starting_price,1+return_list[i,]))
}
total_returns2 <- array(NA,dim=no_sims,dimnames=NULL)
for(i in 1:no_sims){
  total_returns2[i]<-(price_list[i,31]-price_list[i,1])/price_list[i,1]
}

```
## Analysing and visualizing the simulations for HPQ

```{r fig.height=3.5, fig.fullwidth=TRUE}
max2<-which.max(total_returns2)
min2<-which.min(total_returns2)
actual_prices_HPQ <- as.matrix(as.numeric(HPQ$HPQ.Close[(N_days-30):N_days]))
med2<-match(median(total_returns2),total_returns2)

par(mfrow = c(1,2))
lb <- min(price_list[min2,])
ub <- max(price_list[max2,])

#Plotting the min, max, median of the simulated return along with the actual return
plot(0:30, price_list[min2,], type='l', main = 'Simulated price of HPQ',ylab='Price',xlab='Days',ylim=c(lb,ub),col="red")
lines(0:30, price_list[med2,],type='l',col='black')
lines(0:30, price_list[max2,],type='l',col='green')
lines(0:30, actual_prices_HPQ,type='l',col='blue')
grid(lty = 2, col = 'gray', lwd = 1)

hist(total_returns2, main = 'HPQ', xlab = 'Simulated Total Returns', col='blue', freq=TRUE)
actual_return2=(HPQ$HPQ.Close[[N_days]]-HPQ$HPQ.Close[[(N_days-30)]])/HPQ$HPQ.Close[[(N_days-30)]]
abline(v=actual_return2, col = 'red', lwd=2)
abline(v=0, col = 'gray', lty=2, lwd=4)
```
In the plot, Simulations leading to the largest gain and largest loss in total 30-day return are shown in red and green respectively. The median from all performed simulations is shown in black and the actual returns are shown in blue. The histogram shows the distributions of the total returns as well as the actual 30-day return shown in red for comparison. 
## Comparing HPQ stock simulations with the actual return
```{r}
HPQ_expected_return <- mean(total_returns2)
HPQ_return_std <- sd(total_returns2)

summary(total_returns2)
mean(actual_return2)

```
The summary statistics from all Monte Carlo trials had an average of `r mean(total_returns2)`. The actual return is `r actual_return2`. The actual and simulated returns are almost similar for HPQ, thereby the model is accurate for predictions.

## Estimating the probability of gaining or losing more than 10% return during the selected time period for HPQ

```{r}
HPQ_prob_G<-sum(total_returns2 > 0.1)/no_sims
HPQ_prob_L<-sum(total_returns2 < (-0.1))/no_sims

```

As per our simulation, the estimated probability to gain more than 10% in 10 days for HPQ is `r HPQ_prob_G*100`% and the probability to lose more than 10% in 10 days is `r HPQ_prob_L*100`.

## Building a portfolio


We will assume an Initial Investment of $10,000 and allocate our funds between these two stocks based on 3 strategies:

- The entire investment to DIS
- The entire investment to HPQ
- Split the investment evenly between DIS and HPQ

From the previous computations, the cost of DIS is \$`r round(DIS_starting_price,2)`. Our analysis has shown that their expected return is \$`r round(DIS_expected_return, digits=4)` per share with a standard deviation of \$`r round(DIS_return_std, digits=4)`. Shares of HPQ cost \$`r round(HPQ_starting_price,2)` per share, with an expected return of \$`r round(HPQ_expected_return, digits=4)` and a standard deviation of \$`r round(HPQ_return_std, digits=4)`. We will assume that the returns from the two companies are independent. 


```{r}
Initial_Investment <- 10000

X <- round(DIS_expected_return, digits=4) 
st_X <- round(DIS_return_std, digits=4)
num_X <- Initial_Investment/round(DIS_starting_price,2)

Y <- round(HPQ_expected_return, digits=4)
st_Y <- round(HPQ_return_std, digits=4)
num_Y <- Initial_Investment/round(HPQ_starting_price,2)

E_X <- num_X*X
Var_X <- (st_X*num_X)^2

E_Y <- num_Y*Y
Var_Y <- (st_Y*num_Y)^2

E_C <- (num_X/2)*X + (num_Y/2)*Y
Var_C <- (st_X*num_X/2)^2 + (st_Y*num_Y/2)^2

```
- Investing everything in DIS stock will have expected profit of \$`r round(E_X, digits=2)` in 10 days with a standard deviation of `r sqrt(Var_X)`

- Investing everything in HPQ stock will have expected profit of \$`r round(E_Y, digits=2)` in 10 days with standard deviation of `r sqrt(Var_Y)`

- Splitting the investment evenly in both companies will have in 10 days an expected profit of \$`r round(E_C, digits=2)` with standard deviation of `r sqrt(Var_C)`


```{r,echo = FALSE}
# lines for the next part of the Project 1
ClosePrices <- merge(DIS$DIS.Close,HPQ$HPQ.Close)
correlations <- cor(ClosePrices, use = "pairwise.complete.obs")
# plot(as.numeric(ClosePrices[,1]),as.numeric(ClosePrices[,2]),pch=20, xlab='DIS', ylab='HPQ')
```